---
output:
  html_document: default
  pdf_document: default
---
```{r}
library(tidyverse)
library(naniar)
library(scales)
library(ggplot2)
library(forcats)
library(gridExtra)
library(caret)
library(cluster)
library(factoextra)
library(stats)
library(dplyr)
library(ggfortify)
library(GGally)
```

```{r}
#Import data
df = read.csv("IRIS.csv")
```

```{r}
#View data
View(df)
```

```{r}
glimpse(df)
```

```{r}
#First 6 rows
head(df)
```
```{r}
#Check for missing/null values
gg_miss_which(df)
```
```{r}
na.omit(df)
```


```{r}
#Frequency of each species
table(df$species)
```

```{r}
#Number of species different
unique_species <- unique(df$species)
```


```{r}
#Pie chart

colors <- rainbow(length(unique_species))

df %>%
  count(species, sort = TRUE) %>%
  mutate(perc = paste0(round(n / sum(n) * 100.0),'%')) %>%
  ggplot(aes(x = "", y = perc, fill = species)) +
  geom_bar(
    stat = "identity",
    width = 1,
    color = "black",
    size = 1
  ) + 
  coord_polar("y", start = 0) + 
  geom_text(
    aes(y = perc, label = perc), 
    position = position_stack(vjust = 0.5),
    size = 8,
    color = "white",
    fontface = "bold"
  ) + 
  scale_fill_manual(values = colors) + 
  theme_void() + 
  labs(
    title = "Percentage of Species",
    subtitle = "Proportion of Species",
    caption = "Kaggle: Iris Dataset",
    fill = "Species"
  )

```

```{r}
#Iris-setosa Iris-versicolor  Iris-virginica 


plot_multi_histogram <- function(df, feature, label_column) {
    plt <- ggplot(df, aes(x=eval(parse(text=feature)), fill=eval(parse(text=label_column)))) +
    geom_histogram(alpha=0.7, position="identity", aes(y = ..density..), color="black") +
    geom_density(alpha=0.7) +
    labs(x=feature, y = "Density")
    plt + guides(fill=guide_legend(title=label_column))
}

p1 <- plot_multi_histogram(df, "sepal_width", "species")
p2 <- plot_multi_histogram(df, "sepal_length", "species")
p3 <- plot_multi_histogram(df, "petal_length", "species")
p4 <- plot_multi_histogram(df, "petal_width", "species")


grid_plot <- grid.arrange(p1,p2,p3,p4, ncol = 2)
ggsave("histogram_plot.png", plot = grid_plot, width = 10, height = 8, units = "in")

```

```{r}
p1 <- df %>% ggplot(aes(x=species,y=petal_length,fill=species)) +
  geom_boxplot()  

p2 <- df %>% ggplot(aes(x=species,y=sepal_length,fill=species)) +
  geom_boxplot()  

p3 <- df %>% ggplot(aes(x=species,y=sepal_width,fill=species)) +
  geom_boxplot()  

p4 <- df %>% ggplot(aes(x=species,y=petal_width,fill=species)) +
  geom_boxplot()  

grid_plot <- grid.arrange(p1,p2,p3,p4, ncol = 2)
ggsave("box_plot.png", plot = grid_plot, width = 10, height = 8, units = "in")
```

```{r}
p1 <- ggplot(df, aes(x = sepal_length, y = sepal_width, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Sepal Length vs Petal Width",
       x = "Sepal Length",
       y = "Sepal Width")

p2 <- ggplot(df, aes(x = sepal_length, y = petal_length, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Sepal Length vs Petal Length",
       x = "Sepal Length",
       y = "Petal Length")

p3 <- ggplot(df, aes(x = sepal_length, y = petal_width, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Sepal Length vs Petal Width",
       x = "Sepal Length",
       y = "Petal Width")

p4 <- ggplot(df, aes(x = sepal_width, y = petal_length, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Petal Width vs Sepal Length",
       x = "Sepal Width",
       y = "Petal Length")

p5 <- ggplot(df, aes(x = sepal_width, y = petal_width, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Petal Width vs Sepal Length",
       x = "Sepal Width",
       y = "Petal Width")

p6 <- ggplot(df, aes(x = petal_length, y = petal_width, fill = species, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Petal Width vs Sepal Length",
       x = "Petal Length",
       y = "Petal Width")

grid_plot <- grid.arrange(p1,p2,p3,p4, p5, p6, ncol = 2)
ggsave("scatter_plot.png", plot = grid_plot, width = 10, height = 8, units = "in")
```
```{r}
ggpairs(df,mapping = ggplot2::aes(color = df$species),upper = list(continuous = "cor", discrete = "box"),lower = list(continuous = "cor", discrete="dot"))
```

## Performing Kmeans Clustering steps:
Choose the number of clusters k
Select k random points from the data as centroids
Assign all the points to the closest cluster centroid
Recompute the centroids of newly formed clusters
Repeat steps 3 and 4

## Using the elbow method for optimal number of clusters
```{r}
#Finding the optimal number of clusters for k-means classification
x <- scale(df[, 1:4])



fviz_nbclust(x, kmeans, method = "wss")
```


##Gap statistics to choose the number of K
Locate the biggest jump within cluster distance
```{r}
#calculate gap statistic based on number of clusters
gap_stat <- clusGap(x,
                    FUN = kmeans,
                    nstart = 25,
                    K.max = 10,
                    B = 50)

#plot number of clusters vs. gap statistic
print(gap_stat,method="firstmax")
fviz_gap_stat(gap_stat)
```


```{r}
set.seed(123)
#Cluster = 2
k <- 2

kmeans_result <- kmeans(x, centers = k)
str(kmeans_result)
cluster_assignments <- kmeans_result$cluster
cluster_assignments
```
```{r}
autoplot(kmeans_result,x,frame=TRUE)
```

```{r}
fviz_cluster(kmeans_result, x, geom = "point", ellipse.type = "norm")
```
```{r}
aggregate(x, by=list(cluster=kmeans_result$cluster), mean)
```
```{r}
final_result <- cbind(df, cluster = kmeans_result$cluster)
final_result
```







